---
---
---

# Presentation

```{r message=FALSE, warning=FALSE}
library(tidyverse)
install.packages('knitr')
library(knitr)
# Might need to modify next line to load the file from the location where you downloaded it
df <- read_csv("data_for_datathon_2014_2016.csv") # Loads the Data Set into the object df

# Cleaning data
df <- df |> mutate( # Will change the variables in the df data set
  Vaccinations = case_when(is.na(Vaccinations) ~ "NV", .default = Vaccinations), # Reclassifies Missing Data in Vaccination as "N" per code book.
  ReproductiveStatus = case_when(is.na(ReproductiveStatus) ~ "UN", .default = ReproductiveStatus), # Reclassifies Missing Data in Vaccination as "N" per code book.
  BodyCondition = case_when( # Begins to modify BodyCondition
    as.character(BodyCondition) == "1" ~ "emaciated", # Changes 1 to emaciated
    as.character(BodyCondition) == "2" ~ "thin", # Changes 2 to thin
    as.character(BodyCondition) == "3" ~ "healthy wild", # Changes 3 to healthy wild 
    as.character(BodyCondition) == "4" ~ "extra fat reserves", # Changes 4 to Extra Fat Reserve
    as.character(BodyCondition) == "5" ~ "extreme fat reserves" # Changes 5 to Extreme Fat Reserves
  )
)

summary(df)

df <- drop_na(df, AgeClass, Weight, BodyCondition) # Removes the missing values (NA) using the variables AgeClass, Weight, and BodyCondition

nrow(df)

mean_weight<-mean(df$Weight)
sd_weight<-sd(df$Weight)
mean_weight


sd_weight #printing the values

count_df <- table(df$Sex) # Using the 'Sex' variable from the `df` data set, we count the frequencies of each category with the table function and storing it in rs_df.
count_df # Printing out the contents of "rs_df"
prop.table(count_df) # Computing the Proportions of "rs_df"

ggplot(df) + # Setting up the data to create a plot.
  geom_bar(aes(Sex)) + # Creating a bar chart based on the variable "ReproductiveStatus" 
  theme(axis.title = element_text(size = 14), axis.text = element_text(size = 14))


ggplot(df, aes(x = Sex, y = Weight)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Boxplot of Weight by Sex",
    x = "Sex",
    y = "Weight"
  ) + theme_minimal()


weight_summary <- df %>%
  group_by(Sex) %>%
  summarise(
    mean_weight = mean(Weight),
    sd_weight = sd(Weight),
    .groups = "drop"
  )
weight_summary


weight_summary_yearly <- df %>%
  group_by(Sex, SamplingYear) %>%
  summarise(
    mean_weight = mean(Weight, na.rm = TRUE),
    sd_weight = sd(Weight, na.rm = TRUE),
    .groups = "drop"
  ) #group the data by Sex and Year and calculate mean and standard deviation of weight
weight_summary_yearly

ggplot(weight_summary_yearly, aes(x = SamplingYear, y = mean_weight)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = mean_weight - sd_weight,
                    ymax = mean_weight + sd_weight), width = 0.2) +
  facet_wrap(~ Sex) +
  labs(
    title = "Average Weight by Year and Sex",
    y = "Mean Weight",
    x = "Year"
  )

#Santa Rosa Fox population 

SR_df<-filter(df,Island=="SRI")

#San Miguel Fox population
SM_df<-filter(df,Island=="SMI")
```

The fox weight has decreased over the span of 3 years.

Some things to keep in mind when analyzing the fox weight, the majority of the foxes are male (53% to 47%, rounded), which means the weight will inherently run higher as the structure of a male is larger than a female.

This can be seen with the graph on line 39-41. We can see how it effects the weight class by running lines 44-50. By using group_by, we can find that the average weight for females (1.88) kilograms is less than males, (1.99) kilograms.

By looking at the average by year, we can find that the female weight has decreased, 2014, the average weight was 1.95, 2015, 1.85, and 1.86 in 2016, meaning overall a loss of 0.09 kilograms.

Looking at the male population, by looking at the average by year, the male weight has also decreased. In 2014, the average weight was 2.08, in 2015, the average was 1.95, and 1.99 in 2016, meaning an overall a loss of 0.09 kilograms.

**Challenge Questions: 1. Do the foxes on different islands have the same weight? Is there an island where the foxes are significantly heavier than others?**

Foxes on different islands do not have the same weight, as the SRI island has a wider range of weights based on the graph on lines 95-101. SRI has both lighter and heavier foxes

```{r}
ggplot(df, aes(x = Sex, y = Weight, color = Sex)) +
  geom_point() + 
  facet_wrap(~Island) +
  labs(title = "Sex vs Weight Across Different Islands",
       x = "Sex",
       y = "Weight",
       color = "Sex")
```

**2.Do vaccinated foxes have a different weight compared to unvaccinated ones?** It looks like certain vaccines make the foxes thinner, as the graph on lines 104-109 shows there are fewer foxes who received their vaccinations that are on the heavier side.

```{r}

# Do vaccinated foxes have a different weight compared to unvaccinated ones? To answer this, study the relationship between Weight and Vaccinations
ggplot(df, aes(x = Weight, y = Vaccinations, color = Vaccinations)) +
  geom_jitter() + 
  labs(title = "Vaccinations vs Weight Across Different Islands",
       x = "Weight (Kilograms)",
       y = "Vaccines",
       color = "Vaccinations")
```

3\. **The dataset has a variable called BodyCondition. What proportion of foxes are well nourished? Do some islands have a larger proportion of “fat” foxes than others? How does BodyCondition connect with Weight? For the latter, you could look into scatter plots.** If we consider emaciated and thin as not well-nourished, then 0.4702593 are well nourished. With body condition and weight, the heavier the fox, the better the body condition/the more well nourished.

```{r}

#The dataset has a variable called BodyCondition. What proportion of foxes are well nourished? Do some islands have a larger proportion of “fat” foxes than others? How does BodyCondition connect with Weight? For the latter, you could look into scatter plots.

# proportion overall
body_df <- table(df$BodyCondition) 
body_df 
prop.table(body_df)
well_nourished = 0.103202847 + 0.005592272 +  0.361464159 
well_nourished

# Body Condition and Weight
ggplot(df, aes(x = Weight, y = BodyCondition, color=BodyCondition)) +
  geom_point() + 
  facet_wrap(~Island) +
  labs(title = "BodyCondition versus Weight",
       x = "Weight",
       y = "Body Condition",
       color ="Body Condition")
```
